Newly Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated

### 2022-07-03 MScThesisText.tex
almost there

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated},
  pdfauthor={Jiancheng Gu1},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{keywords\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

@@

\title{Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated}
\author{Jiancheng Gu\textsuperscript{1}}
\date{}

@@

\abstract{%
Millions of children worldwide are ``left behind'' when their parents migrate away for work.
Parents' labor migration increases household income but decreases parental care, thereby exerting mixed influences on child development, including educational performance.
To what extent do different arrangements of parental migration affect children's schooling performance in the short term?
Studying the case of China, I obtained a sample of four thousand junior high school students from a nationally representative two-wave panel survey.
I differentiated household migration arrangements of any one parent, of mother only, of father only, and of both parents.
I measured children's educational performance by standardized cognitive test score and academic exam grades.
For causal inference, I built difference-in-differences models with propensity score matching.
Results show that the new left-behind children's performance at school dropped slightly compared to that of non-left-behind peers.
Father-only migration is negatively associated with children's cognitive abilities, while mother-only migration is negatively associated with children's academic grades.
}

@@

Worldwide, millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work. These are the so-called ``left-behind children'' or ``stay-behind children.'' They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children (DeWaard et al., 2018), one-fifth of Bulgarian children (Popova, 2018), and one-fourth of Chinese children were living apart from their migrating parents (NBS et al., 2017). In a 2021 resolution, the Council of Europe's Parliamentary Assembly (2021) expressed concern about ``the scale of this phenomenon and the long-term damage it creates'' (p.~2) and declared that leaving ``millions of labour migrants' children without parental care is a mass violation of human rights'' (p.~1).

How are stay-behind children faring when parents have migrated? Research results are mixed due to the differential effects of parental migration. Consider children's education. According to the review articles by Brauw (2019) and Van Hook and Glick (2020), labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. The inflow of remittances, however, cannot substitute for the loss of parental care. Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance, which merits empirical investigation.

A challenge arises in assessing the causality between parent's migration and children's educational achievement. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. The cause should be on par with random assignment to participants in the analysis, conditional on identification assumptions (Cunningham, 2021).

In this thesis, my research question is: To what extent do different arrangements of parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyzed data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and applied propensity score matching for robustness check. The outcome was measured with children's cognitive test score and academic exam scores. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

\hypertarget{theoretical-framework}{%
\subsection{Theoretical Framework}\label{theoretical-framework}}

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic capital is transferable to human capital, which is the stock of ``skills, talents, health, and expertise'' in people that enhance their productivity (Botev et al., 2019; Goldin, 2016). D. Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational expenditure. In China, Chang et al. (2019) interviewed migrant workers and their family members. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the ``stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships'' (Black et al., 2021, p. 1). Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents (Wen \& Xie, 2019). Hong talked to a group of left-behind children at a rural school (Hong \& Fuller, 2019). These students expressed the utter sense of loneliness ``with no advocate, no support and with no one looking out for them'' (p.~13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p.~14).

Given the two opposing effects, researchers have obtained mixed results regarding how parental migration impacts children's education. Botezat and Pfeiffer (2020) studied the case of Romania and found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen. L. Wang et al. (2019) collected data from junior high school students in central China and found no effect of parental on left-behind children's achievement in mathematical test. Studying four Chinese provinces, Brauw and Giles (2017) showed that as ample migration opportunities appeared in villages, enrollment to senior high school decreased. Nguyen (2016) demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight. The inconsistency can be complicated by factors such as the left-behind children's age and gender, the migrant parent's original domicile and eventual destination, and the arrangement of which parent migrate for how long.

@@

\(H_1\) In the short term, parental migration negatively affect the cognitive development and academic performance of children left-behind.

Researchers have studied whether the impact differ between mother absence and father absence induced by labor migration. Y. Xu et al. (2019) adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had ``little or no association with negative outcomes'' on children's academic and cognitive performances. Only-mother absence, however, showed a ``strong association with negative outcomes'' (p.~1646). Chen et al. (2019) conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was ``negatively associated with children's social competence and academic performance,'' while father migration had no direct effects on the two outcomes (pp.~860-861).

Studies noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, children are more likely to accept fathers' migration than mothers' (Murphy, 2022). Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles. I propose the second hypothesis:

\(H_2\) In the short term, mother's migration negatively affect the cognitive development and academic performance of children left-behind, while father's migration does not show a effect.

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers (NBS, 2022). They form the backbone of essential industries, such as retail, delivery, transportation, and manufacturing. The 2015 one-percent National Population Sample Survey estimated that 68.7 million Chinese children were left behind (NBS et al., 2017). The internal migration stemmed from the neoliberal economic regime.

The neoliberal economic regime sounds at odds with China, a so-called socialist country. Actually, in the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. The regional inequality needs social policy to amend, but China's welfare program is discriminatory -- it segregates the population. Those who reap the greatest benefits are public servants, urban dwellers, and formally employed people. As rural people migrate to the city, they find themselves with limited access to social welfare in the destination. The municipal governments only provide welfare to local residents and expect non-local residents access public services elsewhere (Duckett, 2020; Zhang, 2018).

The institutionalized discrimination against migrant laborers find its roots in China's ``multilevel citizenship'' system. As Vortherms (2021) demonstrated, the constituents of citizenship -- ``membership rules and rights entitlements'' -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through the \emph{hukou}, or household registration system:

In other words, the hukou system binds one's citizenship rights with one's place of registry. In theory, anyone can move out of town freely. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city (Gu, 2022). Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown (Chan, 2019). Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns.

The existing research about China's left-behind children has three major shortcomings. The first is that most literature focuses on children left behind in rural areas while neglecting the urban counterparts. The National Bureau of Statistics stated that the number of urban left-behind children had been growing and expected that the trend would continue (NBS et al., 2017). My sample includes urban left-behind children as well. The second shortcoming lies in sampling and research design. Some studies suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one to a few towns (Y. Liu et al., 2021; Shu, 2021; F. Yang et al., 2022). Others only estimated correlation using cross-sectional data (Akezhuoli et al., 2022; Jin et al., 2020). These studies could have employed causal inference techniques like fixed effect, matching, or inverse probability of treatment weighting. The third shortcoming lies in the measurement of outcome. Chang et al. (2019) and L. Wang et al. (2019) measured students' educational performance with the test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. I included the test results of multiple subjects to measure comprehensive academic achievement.

@@

I obtained the data from the China Education Panel Survey (CEPS) (NSRC, n.d.). It is a nationally representative, school-based survey featuring junior high school students. The survey administrator is the National Survey Research Center at Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties. The follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed 9449 of the seventh grade cohort. I further restricted the sample to students staying with both parents at the original domicile at the baseline survey. To deal with missing values, I adopted list-wise deletion for dependent, independent, and control variables. The final sample size amounted to 4088 .

@@

I explored two categories of educational performance: cognitive and academic abilities. One key dependent variable, the student's cognitive skill, was measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The CEPS research team calculated the raw score based on the Item Response Theory and then standardized the score across the whole national sample with a mean of zero and standard deviation of one (W. Wang \& Li, 2015).

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Newly Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated},
  pdfauthor={Jiancheng Gu1},
  pdflang={en-EN},
  pdfkeywords={labor migration, left-behind children, educational performance},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{labor migration, left-behind children, educational performance\newline\indent Word count: 6,600}
\usepackage{csquotes}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

@@

\title{Newly Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated}
\author{Jiancheng Gu\textsuperscript{1}}
\date{}

@@

\abstract{%
Millions of children worldwide are ``left behind'' at home communities as their parents migrate away for work.
Parents' labor migration increases household income but decreases parental care, thereby exerting mixed influences on children's educational attainment.
To what extent do different arrangements of parental migration affect children's educational performance in the short term?
Studying the case of China, I obtained a sample of four thousand junior high school students from a nationally representative two-wave panel survey.
I differentiated household migration arrangements of any one parent, of mother only, of father only, and of both parents.
I measured children's schooling performance by cognitive test scores and academic exam grades.
For causal inference, I built difference-in-differences models with propensity score matching.
Results show that the newly left-behind children's performance at school dropped slightly compared to that of non-left-behind peers.
Mother-only migration is negatively associated with children's academic grades and cognitive abilities.
Father-only migration is negatively associated with children's cognitive abilities, but its magnitude is greater than mother-only migration.
}

@@

Globally, millions of children reside in their home communities but live apart from their parents as either one or both parents have migrated for work. These are the so-called ``left-behind children'' or ``stay-behind children.'' They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children (DeWaard et al., 2018), one-fifth of Bulgarian children (Popova, 2018), and one-fourth of Chinese children were separated from their migrating parents (NBS et al., 2017). The issue of left-behind children has caught the attention of policymakers and legislators. In a resolution, the Council of Europe's Parliamentary Assembly (2021) expressed concern about ``the scale of this phenomenon and the long-term damage it creates'' and declared that leaving ``millions of labour migrants' children without parental care is a mass violation of human rights''.

How are stay-behind children faring in the absence of parents? Research findings are mixed due to the differential effects of parental migration. Consider children's education. According to the review articles by Brauw (2019) and Van Hook and Glick (2020), labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. To what degree can the inflow of remittances compensate for the loss of parental attention? Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance. This uncertainty merits empirical investigation.

A challenge arises in estimating the effect of parent's migration on children's educational achievement. Experiments can unveil causal relations. It is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. It requires that the cause br on par with random assignment to participants, conditional on identification assumptions (Cunningham, 2021).

In this thesis, my research question is: To what extent do different arrangements of parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyze data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and applied propensity score matching for robustness checks. The outcome was measured with children's cognitive test score and academic exam grade. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers. They form the backbone of essential industries, such as retail, delivery, transportation, and manufacturing. The 2015 one-percent National Population Sample Survey estimated that 68.7 million Chinese children were left behind (NBS, 2022; NBS et al., 2017).

The internal migration stemmed from the neoliberal economic regime. Scholars like Duckett (2020) and Zhan (2020) argued that China conforms to the neoliberal trend, extending the logic of the market to other domains of human life. In the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. Simultaneously, many people have left the countryside for the city. The regional inequality needs social policy to amend. But China's welfare program, as Qian (2021) argued, is fragmented and stratified. Those who reap the greatest benefits are public servants, urban dwellers, and formally employed workers. As rural migrants arrive in the city, they find themselves with limited access to social welfare, such as health care and children's education (Qian, 2021). The welfare system's exclusion of rural migrants means that the state shifts its responsibility to individuals, resembling the neoliberal governmentality (Zhan, 2020).

The institutionalized discrimination against rural migrants finds its roots in China's ``multilevel citizenship'' system. As Vortherms (2021) explained, the constituents of citizenship -- ``membership rules and rights entitlements'' -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through the \emph{hukou}, or household registration system:

In other words, the hukou system binds one's citizenship rights with one's place of registry. In theory, Chinese citizens can migrate freely within the country. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city (Gu, 2022). Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown (Chan, 2019). Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns.

How does China counter the issue of labor migration and left-behind children? According to X. Wang (2020) and Zhang (2018), the government has gradually liberalized the multilevel citizenship scheme. It has introduced residence permit for internal migrants as the basis for social entitlements. First-tier cities like Beijing and Shanghai have brought forth a point-based system to grant hukou to newcomers. Criteria include in-demand talents and skills, educational and professional qualifications, and consecutive years of employment and tax-payment. This parallels how European states differentially incorporate foreign immigrants into national welfare regimes (Zhang, 2018). Another policy response, as stated by Roberts and Hannum (2018), is the centralized boarding schools. The Ministry of Education promotes boarding schools in the rural county-seats. These schools serve as a cost-efficient way to enroll students in surrounding villages and to look after the left-behind children. In reality, such schools ``are often ill-equipped to ensure the healthy upbringing and psychological wellbeing of children'' (p.~6).

Social scientists have researched extensively about China's left-behind children, but the current literature still has three major shortcomings. The first is that most literature focuses on children left behind in rural areas while neglecting the urban counterparts. The National Bureau of Statistics stated that the number of urban left-behind children had been growing and expected that the trend would continue (NBS et al., 2017). My sample includes urban left-behind children as well. The second shortcoming lies in sampling and research design. Some studies suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one to a few towns (Y. Liu et al., 2021; Shu, 2021; F. Yang et al., 2022). Others only estimated correlation using cross-sectional data (Akezhuoli et al., 2022; Huang et al., 2020; Jin et al., 2020). These studies could have employed causal inference techniques like fixed effect, matching, or inverse probability of treatment weighting. The third shortcoming lies in the measurement of outcome. Chang et al. (2019) and L. Wang et al. (2019) measured students' educational performance with the standardized test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. I included the test results of multiple subjects to measure comprehensive academic achievement.

\hypertarget{theory-and-hypotheses}{%
\subsection{Theory and Hypotheses}\label{theory-and-hypotheses}}

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic capital is transferable to human capital, which is the stock of ``skills, talents, health, and expertise'' in people that enhance their productivity (Botev et al., 2019; Goldin, 2016). D. Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational expenditure. Cebotari (2020) developed a fixed effect model with three-wave panel data collected in Ghana. He found that stay-behind children's education benefits from ``the steady inflow of both in-kind and monetary remittances'' contributed by migrant parents. In China, Chang et al. (2019) interviewed migrant workers and their family members. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood. Peng (2018) interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, ``good mothering'' primarily means offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the ``stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships'' (Black et al., 2021, p. 1). Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents (Wen \& Xie, 2019). Hong talked to a group of left-behind children at a rural school (Hong \& Fuller, 2019). These students expressed the utter sense of loneliness ``with no advocate, no support and with no one looking out for them'' (p.~13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p.~14). Similarly, Murphy (2022) talked to left-behind adolescents at a village. Some of them ``embraced the `left-behind' label'' because it captured their need for more parental care (p.~185), ``especially companionship and affection in daily life'' (p.~186).

Given the two opposing effects, researchers have obtained mixed results on how parental migration impacts children's education. Botezat and Pfeiffer (2020) adopted past immigration rate as an instrumental variable. They found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen in Romania. L. Wang et al. (2019) collected data from junior high school students in central China. The researchers' two-way fixed-effect model showed no effect of parental on left-behind children's achievement in mathematical test. Studying four Chinese provinces, Brauw and Giles (2017) showed that as ample migration opportunities appeared in villages, enrollment to senior high school decreased. Nguyen (2016) conducted fixed-effect regression and demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight.

@@

\(H_1\) In the short term, parental migration reduces the educational performance of children left-behind.

\(H_1a\) In the short term, parental migration negatively affects left-behind children's cognitive test score.

\(H_1b\) In the short term, parental migration negatively affects left-behind children's academic exam score.

Researchers have studied whether the impact differ between mother's migration and father's migration. Xu et al. (2019) adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had ``little or no association with negative outcomes'' on children's academic and cognitive performances. Only-mother absence, however, showed a ``strong association with negative outcomes'' (p.~1646). Chen et al. (2019) conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was ``negatively associated with children's social competence and academic performance,'' while father migration had no direct effects on the two outcomes (pp.~860-861).

Scholars like Murphy (2020) noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, fathers' migration is more likely to win children's acceptance than does mothers' migration. Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles (Murphy, 2020). I propose the second hypothesis:

\(H_2\) In the short term, mother's migration more negatively affect the educational performance of children left-behind than do father's migration.

\(H_2a\) In the short term, mother's migration more negatively affect left-behind children's cognitive test score than do father's migration.

\(H_2b\) In the short term, mother's migration more negatively affect left-behind children's academic exam score than do father's migration.

@@

I obtained the data from the China Education Panel Survey (CEPS) (NSRC, n.d.). It is a nationally representative, school-based survey featuring junior high school students. Currently, two waves of data are available. The CEPS team operates under the National Survey Research Center at the Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The baseline survey (wave one) took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties (hereafter `counties'). The follow-up survey (wave two) in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed 9449 of the seventh grade cohort. I further restricted the sample to students living with both parents at the original domicile at the baseline survey. To tackle missing values, I adopted list-wise deletion for the dependent, independent, and control variables. The final sample size amounts to 4062.

@@

I explored two categories of educational performance: cognitive and academic abilities. One key dependent variable, the student's cognitive skill, was measured by a fifteen-minute standardized test administered by the CEPS research team. The test covered three types of reasoning: verbal, visuospatial, and numerical. The baseline and follow-up tests consisted of 20 and 35 items, respectively. The raw score was calculated based on the Item Response Theory (W. Wang \& Li, 2015). I transformed the raw scores to the 100-point scale. I chose the raw score instead of the z-score, as z-standardization fails to document the change of mean between time points, which is also of interest (Moeller, 2015).

Another dependent variable concerned students' academic performance. This was measured by the average scores of mid-term exams in Chinese, mathematics, and English as provided by each school. Exam scores are reliable indicators of learning performance, and are comparable across schools for the same cohort of students. This is related to how junior high schools operate within the compulsory education system in China. According to OECD (2016), China's junior secondary education is a unified system that does not distinguish between academic and vocational tracks. At school, the subjects of Chinese, mathematics, and English are core subjects and others as supplementary subjects. Throughout junior high school, students study the core subjects in preparation for the senior high school entrance exam. Within each cohort in one particular school, teachers use the same syllabus and administer the same exams during a given assessment period. Across cities and regions, junior high schools adopt the educational content and assessment set by the Ministry of Education (OECD, 2016).

I transformed the exam grades through the following steps. I first aligned different grading scales (full mark of 100, 120, and 150 for one subject) with the 100-point scale. After the conversion, I checked for abnormal values lower than zero or higher than one hundred. 30 observations were outliers, accounting for less than one percent of the sample. I retained these observations in the sample and trimmed the subject exam scores. The trimming ensured that all values would be within the 0-100 range. In this way, I can limit the impact of outliers. Finally, I calculated the average score of these three subjects.
